---
title: "191004_HW2"
author: "Gavin Ko"
date: "9/26/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(haven)
library(tidyr)
```

# Problem 1

## A. Mr. Trashwheel Data
(Note. Library section is put in the include=false golbal setting chunk.) 

First of all, data import.

```{r read in data}
Trash_data = 
  read_excel(path = "data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", 
             range = "A2:N338",
             sheet = "Mr. Trash Wheel",
             col_names = TRUE
             ) %>% 
  drop_na(Dumpster) %>% 
  janitor :: clean_names()
  
```

At this place, I've cleared all the missing or unrelated data.
But, some variable names are not clear.
Therefore, I'm going to apply some changes.

```{r rename}
Trash_data = 
  rename(Trash_data, 
         weight = weight_tons,
         volume = volume_cubic_yards,
         )
```

Now it looks better. 
As requested, let's round the sports balls.

```{r rounding sportsballs}
mutate(Trash_data,
       sports_balls = round(sports_balls, digits = 0),
       sports_balls = as.integer(sports_balls)
) %>% 
pull(sports_balls)
```
Now the data is good to go!

## B. Precipitation Data

Let's read in the data for 2017 & 2018 first.
  
```{r read in 2017 & 2018}
Precipitation_2017 = 
  read_excel(path = "data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
             range = "A2:B14",
             sheet = "2017 Precipitation",
             col_names = TRUE) %>% 
  janitor::clean_names() %>% 
  drop_na(total)

Precipitation_2018 = 
  read_excel(path = "data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",
             range = "A2:B14",
             sheet = "2018 Precipitation",
             col_names = TRUE) %>% 
  janitor::clean_names() %>% 
  drop_na(total)
```

Before the combination take place, we need to add the year variable.

```{r adding variable years}
Precipitation_2017 = 
  mutate(Precipitation_2017,
         year = 2017)

Precipitation_2018 = 
  mutate(Precipitation_2018,
         year = 2018)
```

Now, let's try to combine the data.

```{r mutating data}
Precipitation_merged = bind_rows(Precipitation_2017, Precipitation_2018) 
```

This doesn't look really tidy, so let me also change the order.

```{r select}
Precipitation_merged = 
  select(Precipitation_merged, year, everything()) %>% 
  mutate(
  month = as.numeric(month),
  month = month.name[month],)
```

For data summary, I'm going to talk about Mr. Trash Wheel first.
Key variables are the recorded date(`date`), sediment weight(`weight`), volume(`volume`) and homes powered by Mr. Trash Wheel(`homes_powered`). 
After removing missing data, Mr. Trash Wheel dataset has collected record from n =  `r length(pull(Trash_data, dumpster))` dumpsters. 
Among these observations, the average weight of waste was `r round( mean(pull(Trash_data, weight)), 2)` tons, with median at `r median(pull(Trash_data, weight))` tons.
Maximum weight record happened in `r pull(filter(Trash_data, weight == max(weight)), date)`, which recorded a historical `r pull(filter(Trash_data, weight == max(weight)), weight)` tons of waste. In volume scale, the average volume of waste was `r round( mean(pull(Trash_data, volume)), 2)` qubic yrads, with median at `r median(pull(Trash_data, volume))` qubic yards. Maximum volume record happened in `r pull(filter(Trash_data, volume == max(volume)), date)`, which recorded a historical `r pull(filter(Trash_data, volume == max(volume)), volume)` quibic yards of waste. During the data collection period, Mr. Trash Wheel has contributed in total `r sum(pull(Trash_data, weight))` tons, `r sum(pull(Trash_data, volume))` qubic yards of waste, and powered `r as.integer(sum(pull(Trash_data, homes_powered)))` families. Thanks Mr. Trash Wheel!

Secondly, I'm going to talk about Precipitation data. 

After removing missing data, Precipitation dataset has collected record n =  `r length(pull(Precipitation_merged, total))` months of data. 
Among these observations, the mean sediment depth was `r round( mean(pull(Precipitation_merged, total)), 2)` inches, with median at `r median(pull(Precipitation_merged, total))` inches.
Total amount of collected sediment depth during this period was `r sum(pull(Precipitation_merged, total))` inches. Maximum amount of sediment was recorded on `r pull(filter(Precipitation_merged, total == max(total)),month)`, `r pull(filter(Precipitation_merged, total == max(total)),year)` with `r pull(filter(Precipitation_merged, total == max(total)),total)` inches of precipitation.  Specific to precipitation in 2018, the total sediments was `r sum (pull(filter(Precipitation_merged, year == 2018), total))` inches.
Specific to sports balls, the median number of them in 2017 was `r median( pull(filter(Trash_data, year == 2017), sports_balls))`.

# Problem 2
First, read the files.
```{r read files}
Pols_month_data = read.csv("data/pols-month.csv") %>% 
  janitor::clean_names()

Unemployment_data = read.csv("data/unemployment.csv") %>% 
  janitor::clean_names()

Snp_data = read.csv("data/snp.csv") %>% 
  janitor::clean_names()
```
Of course it is not clean enough. Let's deal with pols-month first.

```{r pols month operation}
Pols_month_data = 
  mutate(Pols_month_data,
         mon = as.character(mon)) %>% 
  separate(mon, into = c("year","month","day"), sep = "-") %>%
  mutate(month = month.name[as.numeric(month)],
         president = prez_gop - prez_dem,
         president = factor(president, labels = c("dem","gop","gop"))) %>%
  select(year, month, president, everything(), -day, -prez_dem, -prez_gop)
```

Secondly, time for snp.

```{r snp operation}
Snp_data = 
  separate(Snp_data, date, into = c("month", "day", "year"), sep = "/") %>%
  select(year, month, snp_close = close, -day) %>% 
  mutate(month = as.numeric(month)) %>% 
  arrange(year, month) %>% 
  mutate(month = month.name[month])

```

Lastly, unemployment.

```{r Unemployment operation}
Unemployment_data = 
  pivot_longer(Unemployment_data,
               jan:dec,
               names_to = "month",
               values_to = "unemployment_rate"
               ) %>% 
  mutate(month = month.name[factor(month)],
         year = as.character(year))
```

Last step is to join the data.

```{r snp to pols}
Complete_data = 
  left_join(Pols_month_data, Snp_data, by = c("year","month") ) %>% 
  left_join(Unemployment_data, by = c("year","month")) 
```
First dataset `pols` contains political data at that time: the party of current President (`president`), respective numbers of Governors(`gov_gop, gov_dem`), Senators(`sen_gop, sen_dem`), and Representatives(`rep_gov, rep_dem`). The second dataset `snp` contains the S&P 500 Index at the designated closing month(`snp_close`), which gives a sense of economic status at that time. The last dataset `unemployment` contains unemployment rate in US at that time.
The merged dataset contains `r dim(Complete_data)[1]` months of data with `r dim(Complete_data)[2]` variables recorded. Years ranged from Jan,1947 to Jun, 2015. The NA status in `snp_close` and `unemployment_rate` is owing to the original data collection period. While snp data started from 1950, unemployment rate record started from 1948. 

# Problem 3

First of all, read the data again. During the process, I've also done the lowercases and ethnicity factor adjustment. After that, I've removed the duplicates.

```{r read data and some cleaning}
Pop_baby = read_csv("data/Popular_Baby_Names.csv", col_types = "icccii") %>% 
  janitor::clean_names() %>% 
  mutate(
    gender = factor(str_to_lower(gender)),
    ethnicity = factor(str_to_lower(ethnicity)),
    ethnicity = recode(ethnicity, "asian and paci" = "asian and pacific islander",
                                  "black non hisp" = "black non hispanic",
                                  "white non hisp" = "white non hispanic"),
    childs_first_name = str_to_lower(childs_first_name)) %>% 
  distinct(year_of_birth, gender, ethnicity, childs_first_name, count, rank)
```
Now the data is ready to go!

Dealing with Olivia:
```{r olivia}
Olivia_table = 
  filter(Pop_baby, childs_first_name == "olivia" & gender == "female") %>% 
  select(year_of_birth, ethnicity, rank) %>% 
  pivot_wider(
    names_from = year_of_birth,
    values_from = rank)
Olivia_table

```

Favorite male children names:
```{r best boy name NA}
Male_children_table =
  filter(Pop_baby, rank == 1 & gender == "male") %>% 
  select(year_of_birth, ethnicity, childs_first_name) %>% 
  pivot_wider(
    names_from = year_of_birth,
    values_from = childs_first_name)
Male_children_table

```
Lastly, produce the scatter plot.

```{r filtered data}
  filter(Pop_baby, year_of_birth == 2016 & gender == "male" & ethnicity == "white non hispanic") %>% 
  ggplot(aes(x = rank, y = count)) + geom_point()
```






